Financialisation of Commodity Markets

### 2024-10-30 docs/paper.qmd
minor regime difference test results section edit

Over the second period the pattern essentially reverses. For the twenty-four US commodities over this period, fourteen have higher mean returns during phases of **contango** over **backwardation**. Four of the six GB metals achieved higher mean returns over **contango** relative to **backwardation** phases with the difference being statistically significant for one commodity. The equally weighted US portfolio had a mean return of 15.7% over **backwardation** against 16.2% during phases of **contango**, both statistically significant at the 10% level. The difference was considerably more pronounced for the equally weighted GB metals portfolio with a mean return of 8% during **backwardation** rising to 35.6% during **contango**.

@@

The corresponding pricing performance displays a similar pattern as shown in @tbl-regressions-factors where for each period the corresponding $R^{2}$ values are independently averaged over the entire period as well as over periods of **contango** and **backwardation**. The average $R^{2}$ for the market factor across the entire cross-section of commodity assets, including the GB traded base metals, increases to around 20% during financialisation compared to around 10% over the previous period with the rise driven by the metals sector at large. The sector average for the US traded assets jumps to 31% during financialisation from 7.5% over the past period while for the GB traded assets the corresponding figure increases from less than 2%, or virtually no explanatory power for the (US based) market factor, to almost 12% over the financialisation period. The results for our other risk factor mimicking portfolios show a similar pattern and provide a deeper insights into co-movement dynamics in this context. All three factors---CHP, open interest and term structure---performed poorly in the first period, indicating that the commodities with extreme factor attributes (high or low CHP for example) were uncorrelated with the rest of the commodity cross-section, and the overall performance consistent across periods of **contango** and **backwardation**. The onset of financialisation leads to a substantial improvement in the performance of the CHP factor with its average $R^{2}$, albeit still small at 5.4%, almost tripling from the past to the financialisation period in contrast to the other factors that show weaker dynamics with a slight decrease and a much smaller increase for the open interest and term structure factors respectively. Here again the increase in pricing performance for the CHP factor is driven by the metals sector, particularly the precious sub-sector with the average $R^{2}$ for the group jumping to almost 26% during financialisation from 1.2% over the past period. The corresponding figures of 22% and 1.1% for the entire US metals sector stand in sharp contrast with the agricultural sector, which exhibits no notable change, and the energy sector, where pricing performance drops substantially from 9.2% down to 1.6%. The improvement in the CHP factor's pricing performance remarkably comes predominantly from the long leg (low CHP assets: Keynesian "normal backwardation") of the factor^[Results available from the authors upon request.] with the corresponding mimicking portfolio's performance itself driven by the metals, the precious sub-sector in particular. The average $R^{2}$ for the group jumps to 55% during financialisation from less than 12% over the past period while the short leg sees a far smaller increase, up to 8.9% from 3.3% over the past period. Interestingly, although the term structure factor itself only sees a very modest increase, its short leg (low roll-yield assets: Working's backwardation) shows a sizeable rise in pricing performance, yet again driven the metals sector, with the average $R^{2}$ for the group's US traded assets jumping to 28.2% during financialisation from 6.2% over the past period. The pattern of results for both factors extends to the GB traded assets and suggest a convergence of Keneysian and Working's term structure perspective over the financialisation period.   
Taken together these results emphasise the global nature of the impact of financialisation on the metals complex and provide further evidence that global metals co-movement could be detected by US based factor models that combine the Keynesian and Working’s term structure paradigms. There seems thus to be information in the factor constructed from hedging pressure based on the Keynesian paradigm relevant to the term structure factor built on the Working paradigm and these findings therefore appear to complement the results of @kang_tale_2020 who find that hedging pressure both conveys information about liquidity provision in the short term (Working paradigm) and risk transfer in the longer term (Keynesian paradigm).

>>>>

Over the second period the pattern essentially reverses. For the twenty-four US commodities over this period, fourteen have higher mean returns during phases of **contango** over **backwardation**. Four of the six GB metals achieved higher mean returns over **contango** relative to **backwardation** phases with the difference being statistically significant at the 5% level for one commodity (Nickel-primary (XLME)). The equally weighted US portfolio had a mean return of 15.7% over **backwardation** against 16.2% during phases of **contango**, both statistically significant at the 10% level. The difference was considerably more pronounced for the equally weighted GB metals portfolio with a mean return of 8% during **backwardation** rising to 35.6% during **contango**.

@@

The corresponding pricing performance displays a similar pattern as shown in @tbl-regressions-factors where for each period the corresponding $R^{2}$ values are independently averaged over the entire period as well as over periods of **contango** and **backwardation**. The average $R^{2}$ for the market factor across the entire cross-section of commodity assets, including the GB traded base metals, increases to around 20% during financialisation compared to around 10% over the previous period with the rise driven by the metals sector at large. The sector average for the US traded assets jumps to 31% during financialisation from 7.5% over the past period while for the GB traded assets the corresponding figure increases from less than 2%, or virtually no explanatory power for the (US based) market factor, to almost 12% over the financialisation period. The results for our other risk factor mimicking portfolios show a similar pattern and provide a deeper insights into co-movement dynamics in this context. All three factors---CHP, open interest and term structure---performed poorly in the first period, indicating that the commodities with extreme factor attributes (high or low CHP for example) were uncorrelated with the rest of the commodity cross-section, and the overall performance consistent across periods of **contango** and **backwardation**. The onset of financialisation leads to a substantial improvement in the performance of the CHP factor with its average $R^{2}$, albeit still small at 5.4%, almost tripling from the past to the financialisation period in contrast to the other factors that show weaker dynamics with a slight decrease and a much smaller increase for the open interest and term structure factors respectively. Here again the increase in pricing performance for the CHP factor is driven by the metals sector, particularly the precious sub-sector with the average $R^{2}$ for the group jumping to almost 26% during financialisation from 1.2% over the past period. The corresponding figures of 22% and 1.1% for the entire US metals sector stand in sharp contrast with the agricultural sector, which exhibits no notable change, and the energy sector, where pricing performance drops substantially from 9.2% down to 1.6%. The improvement in the CHP factor's pricing performance remarkably comes predominantly from the long leg (low CHP assets: Keynesian "normal backwardation") of the factor^[Results available from the authors upon request.] with the corresponding mimicking portfolio's performance itself driven by the metals, the precious sub-sector in particular. Interestingly, although the term structure factor itself only sees a very modest increase, its short leg (low roll-yield assets: Working's backwardation) shows a sizeable rise in pricing performance, yet again driven the metals sector. The pattern of results for both factors extends to the GB traded assets and suggest a convergence of Keneysian and Working's term structure perspective over the financialisation period.   
Taken together these results emphasise the global nature of the impact of financialisation on the metals complex suggest that there is information in the factor constructed from hedging pressure based on the Keynesian paradigm relevant to the term structure factor built on the Working paradigm. These findings therefore appear to complement the results of @kang_tale_2020 who find that hedging pressure both conveys information about liquidity provision in the short term (Working paradigm) and risk transfer in the longer term (Keynesian paradigm).
